---
title: "Get the OSM network"
author: "Carole Voulgaris"
date: "9/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required libraries

```{r}
library(tidyverse)
library(tidytransit)
library(sf)
library(tidycensus)
library(r5r)
library(osmdata)
```
You can modify this code to get the osm street network corresponding to a particular transit network.

# Get the extent of the transit netw

1. Get bounding box for the transit network

```{r}
LA_bus <- read_gtfs("Networks/LA/1Feb/GTFS.zip")

LA_stops <- st_as_sf(LA_bus$stops, coords = c("stop_lon", "stop_lat"), crs = "WGS84")
LA_extent <- st_bbox(LA_stops)

opq(bbox = LA_extent, timeout = 120) %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'Networks/LA/1Feb/osmdata_network.osm')

```

At this point, you'll need to use [osmconvert](https://wiki.openstreetmap.org/wiki/Osmconvert){target="_blank"} to convert the osm file to a pbf file. 

I've saved a copy of the executable to the "Networks" folder. Copy it into the `Networks/LA/1Feb/` folder, and run it from your computer (you're using a PC, right?)

