---
title: "Service Changes"
author: "Carole Voulgaris"
date: "9/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required libraries

```{r, message = FALSE}
library(tidyverse)
library(tidytransit)
library(knitr)
```

# Los Angeles

Average weekday frequency before COVID

```{r, message=FALSE}
LA_pre_covid <- read_gtfs("Networks/LA/1Feb/GTFS.zip")

LA_pre_covid_calendar <- LA_pre_covid$calendar

LA_pre_covid_trips <- LA_pre_covid$trips

# filter trips for typical weekday
LA_pre_covid_calendar_wkdy <- LA_pre_covid_calendar %>%
  filter(monday == 1 &
           tuesday == 1 &
           wednesday == 1 &
           thursday == 1 &
           friday == 1)

LA_pre_covid_trips_wkdy <- LA_pre_covid_trips %>%
  left_join(LA_pre_covid_calendar_wkdy) %>%
  filter(!is.na(monday))

LA_pre_covid_wkdy_freq <- LA_pre_covid_trips_wkdy %>%
  group_by(route_id) %>%
  summarise(daily_trips_pre = n())

```

Average Weekday frequency during COVID

```{r, message=FALSE}
LA_covid <- read_gtfs("Networks/LA/31May/GTFS.zip")

LA_covid_calendar <- LA_covid$calendar

LA_covid_trips <- LA_covid$trips

# filter trips for typical weekday
LA_covid_calendar_wkdy <- LA_covid_calendar %>%
  filter(monday == 1 &
           tuesday == 1 &
           wednesday == 1 &
           thursday == 1 &
           friday == 1)

LA_covid_trips_wkdy <- LA_covid_trips %>%
  left_join(LA_covid_calendar_wkdy) %>%
  filter(!is.na(monday))

LA_covid_wkdy_freq <- LA_covid_trips_wkdy %>%
  group_by(route_id) %>%
  summarise(daily_trips_covid = n())

```

Calculate changes by line.

```{r, message=FALSE, warning=FALSE}
compare_freq_LA <- full_join(LA_pre_covid_wkdy_freq, LA_covid_wkdy_freq) %>%
  mutate(pct_chng = (daily_trips_covid - daily_trips_pre) / daily_trips_pre) 

LA_hist <- ggplot(compare_freq_LA) +
  geom_histogram(aes(x = pct_chng), color = "white") +
  scale_x_continuous(name = "Percent change in weekday frequency, Feb - May, 2020",
                     breaks = breaks <- seq(-0.7, 0.5, by=0.1),
                     labels = paste(prettyNum(breaks*100, digits=0), "%", sep="")) +
  scale_y_continuous(name = "Number of lines",
                     limits = c(0,25),
                     breaks = seq(0,24, by=2)) +
  theme_bw()

LA_hist
```

Present the results in a neat little table.

```{r}
LA_summary <- tibble(City = "Los Angeles",
                     `Lines with no frequency change` = 
                       table(compare_freq_LA$pct_chng == 0)["TRUE"],
                     `Lines with reduced frequency` =
                       table(compare_freq_LA$pct_chng < 0)["TRUE"],
                     `Lines with increased frequency` = 
                       table(compare_freq_LA$pct_chng > 0)["TRUE"],
                     `Maximum frequency reduction` = 
                       -1 * min(compare_freq_LA$pct_chng, na.rm=TRUE),
                     `Maximum frequency increase` =
                       max(compare_freq_LA$pct_chng, na.rm=TRUE))

kable(LA_summary)
```

# Dallas

Average weekday frequency before COVID

```{r, message=FALSE}
Dallas_pre_covid <- read_gtfs("Networks/Dallas/1Feb/GTFS.zip")

Dallas_pre_covid_calendar <- Dallas_pre_covid$calendar

Dallas_pre_covid_trips <- Dallas_pre_covid$trips %>%
  left_join(Dallas_pre_covid$routes) %>%
  select(service_id, trip_id, route_short_name)

# filter trips for typical weekday
Dallas_pre_covid_calendar_wkdy <- Dallas_pre_covid_calendar %>%
  filter(monday == 1 &
           tuesday == 1 &
           wednesday == 1 &
           thursday == 1 &
           friday == 1)

Dallas_pre_covid_trips_wkdy <- Dallas_pre_covid_trips %>%
  left_join(Dallas_pre_covid_calendar_wkdy) %>%
  filter(!is.na(monday))

Dallas_pre_covid_wkdy_freq <- Dallas_pre_covid_trips_wkdy %>%
  group_by(route_short_name) %>%
  summarise(daily_trips_pre = n())

```

Average Weekday frequency during COVID

No Friday service?

```{r, message=FALSE}
Dallas_covid <- read_gtfs("Networks/Dallas/31May/GTFS.zip")

Dallas_covid_calendar <- Dallas_covid$calendar

Dallas_covid_trips <- Dallas_covid$trips %>%
  left_join(Dallas_covid$routes) %>%
  select(service_id, trip_id, route_short_name)

# filter trips for typical weekday
Dallas_covid_calendar_wkdy <- Dallas_covid_calendar %>%
  filter(monday == 1 &
           tuesday == 1 &
           wednesday == 1 &
           thursday == 1)

Dallas_covid_trips_wkdy <- Dallas_covid_trips %>%
  left_join(Dallas_covid_calendar_wkdy) %>%
  filter(!is.na(monday))

Dallas_covid_wkdy_freq <- Dallas_covid_trips_wkdy %>%
  group_by(route_short_name) %>%
  summarise(daily_trips_covid = n())

```

Calculate changes by line.

```{r, message=FALSE, warning=FALSE}
compare_freq_Dallas <- full_join(Dallas_pre_covid_wkdy_freq, Dallas_covid_wkdy_freq) %>%
  mutate(pct_chng = (daily_trips_covid - daily_trips_pre) / daily_trips_pre) 

Dallas_hist <- ggplot(compare_freq_Dallas) +
  geom_histogram(aes(x = pct_chng), color = "white") +
  scale_x_continuous(name = "Percent change in weekday frequency, Feb - May, 2020",
                     breaks = breaks <- seq(-0.7, 0.5, by=0.1),
                     labels = paste(prettyNum(breaks*100, digits=0), "%", sep="")) +
  scale_y_continuous(name = "Number of lines",
                     limits = c(0,25),
                     breaks = seq(0,24, by=2)) +
  theme_bw()

Dallas_hist
```

Present the results in a neat little table.

```{r}
Dallas_summary <- tibble(City = "Dallas",
                     `Lines with no frequency change` = 
                       table(compare_freq_Dallas$pct_chng == 0)["TRUE"],
                     `Lines with reduced frequency` =
                       table(compare_freq_Dallas$pct_chng < 0)["TRUE"],
                     `Lines with increased frequency` = 
                       table(compare_freq_Dallas$pct_chng > 0)["TRUE"],
                     `Maximum frequency reduction` = 
                       -1 * min(compare_freq_Dallas$pct_chng, na.rm=TRUE),
                     `Maximum frequency increase` =
                       max(compare_freq_Dallas$pct_chng, na.rm=TRUE))

city_summary = rbind(LA_summary, Dallas_summary)

kable(city_summary)
```

# New York

Average weekday frequency before COVID

```{r, message=FALSE}
NY_pre_covid <- read_gtfs("Networks/New York/1Feb/GTFS.zip")

Dallas_pre_covid_calendar <- Dallas_pre_covid$calendar

Dallas_pre_covid_trips <- Dallas_pre_covid$trips %>%
  left_join(Dallas_pre_covid$routes) %>%
  select(service_id, trip_id, route_short_name)

# filter trips for typical weekday
Dallas_pre_covid_calendar_wkdy <- Dallas_pre_covid_calendar %>%
  filter(monday == 1 &
           tuesday == 1 &
           wednesday == 1 &
           thursday == 1 &
           friday == 1)

Dallas_pre_covid_trips_wkdy <- Dallas_pre_covid_trips %>%
  left_join(Dallas_pre_covid_calendar_wkdy) %>%
  filter(!is.na(monday))

Dallas_pre_covid_wkdy_freq <- Dallas_pre_covid_trips_wkdy %>%
  group_by(route_short_name) %>%
  summarise(daily_trips_pre = n())

```

Average Weekday frequency during COVID

No Friday service?

```{r, message=FALSE}
Dallas_covid <- read_gtfs("Networks/Dallas/31May/GTFS.zip")

Dallas_covid_calendar <- Dallas_covid$calendar

Dallas_covid_trips <- Dallas_covid$trips %>%
  left_join(Dallas_covid$routes) %>%
  select(service_id, trip_id, route_short_name)

# filter trips for typical weekday
Dallas_covid_calendar_wkdy <- Dallas_covid_calendar %>%
  filter(monday == 1 &
           tuesday == 1 &
           wednesday == 1 &
           thursday == 1)

Dallas_covid_trips_wkdy <- Dallas_covid_trips %>%
  left_join(Dallas_covid_calendar_wkdy) %>%
  filter(!is.na(monday))

Dallas_covid_wkdy_freq <- Dallas_covid_trips_wkdy %>%
  group_by(route_short_name) %>%
  summarise(daily_trips_covid = n())

```

Calculate changes by line.

```{r, message=FALSE, warning=FALSE}
compare_freq_Dallas <- full_join(Dallas_pre_covid_wkdy_freq, Dallas_covid_wkdy_freq) %>%
  mutate(pct_chng = (daily_trips_covid - daily_trips_pre) / daily_trips_pre) 

Dallas_hist <- ggplot(compare_freq_Dallas) +
  geom_histogram(aes(x = pct_chng), color = "white") +
  scale_x_continuous(name = "Percent change in weekday frequency, Feb - May, 2020",
                     breaks = breaks <- seq(-0.7, 0.5, by=0.1),
                     labels = paste(prettyNum(breaks*100, digits=0), "%", sep="")) +
  scale_y_continuous(name = "Number of lines",
                     limits = c(0,25),
                     breaks = seq(0,24, by=2)) +
  theme_bw()

Dallas_hist
```

Present the results in a neat little table.

```{r}
Dallas_summary <- tibble(City = "Dallas",
                     `Lines with no frequency change` = 
                       table(compare_freq_Dallas$pct_chng == 0)["TRUE"],
                     `Lines with reduced frequency` =
                       table(compare_freq_Dallas$pct_chng < 0)["TRUE"],
                     `Lines with increased frequency` = 
                       table(compare_freq_Dallas$pct_chng > 0)["TRUE"],
                     `Maximum frequency reduction` = 
                       -1 * min(compare_freq_Dallas$pct_chng, na.rm=TRUE),
                     `Maximum frequency increase` =
                       max(compare_freq_Dallas$pct_chng, na.rm=TRUE))

city_summary = rbind(LA_summary, Dallas_summary)

kable(city_summary)
```



# Chicago




